---
title: "Homework4"
author: "Peter Jenkins"
date: "3/15/2021"
output: html_document
---
# Homework
#1. Fit a Poisson regression model that assumes expected count is an interactive function of variables x1
#and x2.
```{r}
dat <- read.csv('Homework 4 Data.csv')
head(dat)
fit <- glm(y ~ x1 * x2, family = poisson, data = dat)

summary(fit)

```


#2. Interpret the effect of variable x1 on the expected count when x2 is fixed at level "b". Verify your interpretation in R.
```{r}
b <- coef(fit)
b
# log expected count at x1 = 1
x1.1 <- b[1] + b[2] * 1 + b[3] * 1+ b[4]
x1.1

# Log expected count at x1 = 2 
x1.2 <- b[1] + b[2] * 2 +  b[3] * 1 + b[4]
x1.2

log(exp(x1.2) / exp(x1.1))
b[2]

# expressed as count change
exp(x1.2) / exp(x1.1)

exp(b[2])

```


The log expected count decreases by -0.0994 for each unit increase of x1 when x2 is fixed at level 'b'. 
The expected count increased by 0.905 for each unit increase of x1 when x2 is fixed at level 'b'. 

#3. Interpret the effect of variable x2 on the expected count when x1 is fixed at 1. Verify your interpretation in R.
```{r}

# log expected count at x1 = 1
x2b <- b[1] + b[2] * 1 + b[3] * 1+ b[4]
x2b
# Log expected count at x1 = 2 
x2a <- b[1] + b[2] * 1  + b[4]
x2a

log(exp(x2b) / exp(x2a))
b[3]

# expressed as count change
exp(x2b) / exp(x2a)

exp(b[3])

```

The change in expected count when x1 = 1 is 0.351 between groups a and b. 


#4. Predict the expected count, ± 95% confidence intervals, over the observed range of values of x1, assuming x2 is fixed at level "b".
```{r}

nd <- data.frame(
  x1 = seq(min(dat$x1), max(dat$x1), length.out = 100),
  x2 = factor(x = rep('b', times = 100),
                   levels = c('a', 'b'))
  
)
prd <- predict.glm(object = fit, newdata = nd, type = 'link', se.fit = T)

low <- exp(prd$fit + qnorm(0.025) * prd$se.fit)# quantile related to 2.5% on one side of tail for 5% A
# or low <- exp(prd$fit - qnorm(0.975) * pred$se.fit)
high <- exp(prd$fit - qnorm(0.025) * prd$se.fit)

# plotting predicted values
plot(y = exp(prd$fit), x = nd$x1, xlab = 'x1',
     ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(low), max(high)), type = 'l')
lines(x = nd$x1, y = low, lty = 2)
lines(x = nd$x1, y = high, lty = 2)


```

#5. Predict the expected count, ± 95% confidence intervals, of levels "a" and "b", assuming x1 is fixed at it’s mean.
```{r}

nd <- data.frame(
  x1 = rep(mean(dat$x1), times = 1),
  x2 = factor(x = rep(c('a','b'), times = 1),
              levels = c('a', 'b'))
  
)
prd <- predict.glm(object = fit, newdata = nd, type = 'link', se.fit = T)

low <- exp(prd$fit + qnorm(0.025) * prd$se.fit)# quantile related to 2.5% on one side of tail for 5% A
# or low <- exp(prd$fit - qnorm(0.975) * pred$se.fit)
high <- exp(prd$fit - qnorm(0.025) * prd$se.fit)

expA <- exp(prd$fit[1]) 
expACI <- c(low[1], high[1])

```
