---
title: "Homework 1"
author: "Peter Jenkins"
date: "February 7, 2021"
output: html_document
---

Question 1. 


Loading the data and viewing header. 
```{r}
#install.packages("ggplot2")
library("ggplot2")

beach <- read.csv("beach.csv")
head(beach)
```

2. Fit a model that includes interactions between Year and BeachID. Report estimates of all coefficients

```{r}
fit <-lm(formula = OpenBeach ~ Year * BeachID , data = beach)
summary(fit)
```
3. Predict the beach area at all three beaches in 1992

```{r}
predict(fit, data.frame(Year=1992, BeachID = c("A","B","C")))

```


4. Plot the predicted area of each beach across all years

```{r}

predall <- predict(fit, beach)
names(predall) <- beach$BeachID
predall

```
 
 
 
 4. Plot the predicted area of each beach across all years
```{r}
plot(beach$Year,predall, pch= beach$BeachID)
```

 5. Interpret the rate of beach area change at each of the 3 beaches
```{r}
beacha <- subset(beach, beach$BeachID == "A") 
beachb <- subset(beach, beach$BeachID == "B")
beachc <- subset(beach, beach$BeachID== "C")



fit <-lm(formula = OpenBeach ~ Year  , data = beacha)
summary(fit)

fit <-lm(formula = OpenBeach ~ Year  , data = beachb)
summary(fit)

fit <-lm(formula = OpenBeach ~ Year  , data = beachc)
summary(fit)

```
 Beach A decreases by 18.365 ha each year. 
 
Beach B decreases by -1.3221 ha each year. 

Beach C decreases by -3.994 ha each year. 
 
 Please hand in both the html document and the .Rmd document to GitHub.