---
title: "Homework 2"
author: "Peter Jenkins"
date: "February 11, 2021"
output: html_document
---

1. Calculate the residuals. Use the summary() function applied to your numeric vector of residuals to
verify that this matches the Residuals: quantiles reported from the summary of your fitted model.
Plot a histogram of your residuals. Do they appear reasonably Gaussian? (we will go into detail on
formal model checking later in the semester)

```{r}
#install.packages("ggplot2")
library("ggplot2")

beach <- read.csv("beach.csv")
head(beach)
fit <-lm(formula = OpenBeach ~ Year * BeachID , data = beach)
summary(fit)


# first calculating dummy variables for levels b and c
beach$b <- ifelse(beach$BeachID == 'B', 1, 0)
beach$c <- ifelse(beach$BeachID == 'C', 1, 0)

# storing coefficient for quick retrieval
b <- coef(fit)
b
# predicting new values for y (manually calculating y) 
prd <- b[1] +               # intercept value 
  b[2] * beach$Year  +           #year 
 b[3] * beach$b     +        # b dummy var
 b[4] * beach$c      +        # c dummy variable 
 b[5] * beach$Year * beach$b   +           # interaction year and b 
 b[6] * beach$Year * beach$c    # interaction betweenyear and c 

prd 

# calculating residuals ( real - predited values)
res <- beach$OpenBeach - prd
summary(res)


hist(res)
```
 Yes based off of the Histogram above the residuals are close enough a normal distribution. 

2. Calculate test statistics for your regression coefficients. Verify by comparing to test statistics reported from model output.
```{r}
######
## for intercept
#####

b_0 <- coef(fit)[1]
#b_0
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['(Intercept)','Std. Error']
#s_b
# null hypothesis
B_0 <- 0

# test stat   coefficient - null hypothesis (in this case 0) / standard error  
t1 <- (b_0 - B_0)/ s_b
t1


##########
### returns fitted model slope coef
##########  
b_1 <- coef(fit)[2]
#b_1
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['Year','Std. Error']
#s_b
# null hypothesis
B_1 <- 0

# test stat
t2 <- (b_1 - B_1)/ s_b
t2
#####

######
## for Beach B
#####

b_3 <- coef(fit)[3]
#b_0
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['BeachIDB','Std. Error']
#s_b
# null hypothesis
B_0 <- 0

# test stat   coefficient - null hypothesis (in this case 0) / standard error  
t3 <- (b_3 - B_0)/ s_b
t3
#####


######
## for Beach C
#####

b_4 <- coef(fit)[4]
#b_0
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['BeachIDC','Std. Error']
#s_b
# null hypothesis
B_0 <- 0

# test stat   coefficient - null hypothesis (in this case 0) / standard error  
t4 <- (b_4 - B_0)/ s_b
t4
#####

######
## for Beach B Interaction with Year
#####

b_5 <- coef(fit)[5]
#b_0
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['Year:BeachIDB','Std. Error']
#s_b
# null hypothesis
B_0 <- 0

# test stat   coefficient - null hypothesis (in this case 0) / standard error  
t5 <- (b_5 - B_0)/ s_b
t5
#####

######
## for Beach C Interaction with Year
#####

b_6 <- coef(fit)[6]
#b_0
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['Year:BeachIDC','Std. Error']
#s_b
# null hypothesis
B_0 <- 0

# test stat   coefficient - null hypothesis (in this case 0) / standard error  
t6 <- (b_6 - B_0)/ s_b
t6
#####

########
```
3. Calculate p-values for your regression coefficients. Verify by comparing to p-values reported from model
output. What are the associated null hypotheses? Do you reject or fail to reject these null hypotheses?
Note: in the lecture slides I reported test statistics are t-distributed with n − 2 degrees of freedom.
More generally, these test statistics are t-distributed with n − k degrees of freedom, where k is the
number of regression coefficients in your linear model. (5 points)
```{r}
# under assumptions of null hypothesis the test stat is t distributed with n-2 degree freedom 
rdf <- nrow(beach) - 6  # 62 - 6 number of observations minus total number of coefficient

 tval <- rbind(t1, t2, t3, t4, t5,t6)

#####
#  P value 
#####
# prob of observation larger or smaller than test statistic ************ homework 
# two tailed
for( i in tval) {

 pval <- 2 * pt(abs(i), df = rdf, lower.tail = FALSE)
 print(pval)}  # abs allow for always positive. -1 in first part puts you in lower and left hand side...  1- abs gives upper side and to the right hand side 
#####
summary(fit)

```
The null hypothesis are all that there is no impact on the area of open beach. H0 = 0  Ha  does not = 0 .  We rejected the null hypothesis because of our small pvalues. 





4. Select a single regression coefficient (your choice) and devise a null hypothesis that is different from
the default in lm(). Report the test statistics, your p-value, and whether you reject or fail to reject
your null hypothesis. (5 points)

```{r}
##########
### returns fitted model slope coef or impact of year on beach
# null hypothesis is that we lost an average of 10 ha of beach based on previous study of sea level rise
##########  
b_1 <- coef(fit)[2]
#b_1
# estimate standard error 
s_b <- summary(fit)[["coefficients"]]['Year','Std. Error']
#s_b
# null hypothesis
B_1 <- -10

# test stat
t2 <- (b_1 - B_1)/ s_b
t2
#####

rdf <- nrow(beach) - 6  # 62 - 6 number of observations minus total number of coefficient

 tval <- rbind(t1, t2, t3, t4, t5,t6)

#####
#  P value 
#####
# prob of observation larger or smaller than test statistic ************ homework 
# two tailed
 pval <- 2 * pt(abs(t2), df = rdf, lower.tail = FALSE)
 pval
   # abs allow for always positive. -1 in first part puts you in lower and left hand side...  1- abs gives upper side and to the right hand side 
#####


```
 Our null hypothesis was that B_1= -10 .  
 
Test statistic = -6.58
Pval = 1.72 e-08 

We reject our null hypothesis. Which means we are losing more beach than 10 ha a year. 
 
 
 
 
 5. Interpret output of your fitted model. Tell me how beach area does (or does not change) through time
at each of the 3 beaches.
```{r}
predall <- predict(fit, beach)
names(predall) <- beach$BeachID

plot(beach$Year,predall, pch= beach$BeachID)

beacha <- subset(beach, beach$BeachID == "A") 
beachb <- subset(beach, beach$BeachID == "B")
beachc <- subset(beach, beach$BeachID== "C")



fit <-lm(formula = OpenBeach ~ Year  , data = beacha)
summary(fit)

fit <-lm(formula = OpenBeach ~ Year  , data = beachb)
summary(fit)

fit <-lm(formula = OpenBeach ~ Year  , data = beachc)
summary(fit)

```
 Beach A decreases by 18.365 ha each year. 
 
Beach B decreases by -1.3221 ha each year. 

Beach C decreases by -3.994 ha each year. 
 



 Please hand in both the html document and the .Rmd document to GitHub.