---
title: "HW7"
author: "PeterJenkins"
date: "April 19, 2021"
output: html_document
---
#1. Load data and place into an unmarkedFrameOccu object count data
```{r}

y <- read.csv('Bobcat.csv')
bob_mat <- as.matrix(y)
# detection covariates
p_covs <- read.csv('p covariates.csv')
det_covs <- list(
  people = p_covs)
  


# site-level covariates
site_covs <- read.csv('psi covariates.csv')

library(unmarked)
# formatting for pcount()
nmix_data <-
  unmarkedFrameOccu(y = bob_mat, # detection / non-detection
                      siteCovs = site_covs, # site-level covs
                      obsCovs = det_covs) # detection covariates


```




#2. Fit the following candidate set of models:
  
 # detection model     occupancy model
#intercept-only       intercept-only
#people                intercept-only
#intercept-only           disturbance
#people                 disturbance

```{r}

fit_int <- occu(~ 1 ~ 1, data = nmix_data)
summary(fit_int)


fit_pInt <- occu(~ people ~ 1, data = nmix_data)
summary(fit_pInt)

fit_IntDist <- occu(~ 1 ~ Dist_5km, data = nmix_data)
summary(fit_IntDist)


fit_PDist <- occu(~ people ~ Dist_5km, data = nmix_data)
summary(fit_PDist)

```






#3. Perform model selection with AIC. What is your top model? How do you know? Is there model selection uncertainty?
```{r}

#install.packages("AICcmodavg")
library(AICcmodavg)

cand.set <- list(
  P1 = fit_int, P2 = fit_pInt, P3 = fit_IntDist, P4 = fit_PDist)

mods <- aictab(cand.set = cand.set, second.ord = F)
head(mods)



```
The model with the Intercept and Distance is the best model because it has the lowest AIC value, however People and Distance also has support at the delta AIC is within 2 of the top model. 
  
  
#4. Average both the effect of people on detection, and disturbance on occupancy, over all models. Report model-averaged slope coefficients and 95% confidence intervals.
```{r}

avg_type_p <- modavgShrink(cand.set = cand.set,
                           parm = 'people',
                           second.ord = F,
                           parm.type = 'detect')

avg_type_p$Mod.avg.beta


avg_type_p$Lower.CL
avg_type_p$Upper.CL



avg_type_dist <- modavgShrink(cand.set = cand.set,
                           parm = 'Dist_5km',
                           second.ord = F,
                           parm.type = 'psi')

avg_type_dist$Mod.avg.beta


avg_type_dist$Lower.CL
avg_type_dist$Upper.CL



```




#5. Obtain and plot model-averaged predictions of occupancy probability and detection probability. Average over all models, and make predictions over the observed range of each variable.
```{r}
apply(X = p_covs[,-1], MARGIN = 2, FUN = max, na.rm = TRUE) # determing max number of people observed 
apply(X = p_covs[,-1], MARGIN = 2, FUN = min, na.rm = TRUE) # determing max number of people observed



new_dat <- data.frame(
  people = seq(min(0), max(5.60), length.out = 100),
  Dist_5km = seq(min(site_covs$Dist_5km), max(site_covs$Dist_5km), length.out = 100)
  
)

  



# calculating model averaged predictions occupancy 
avg_prd <- modavgPred(cand.set = cand.set,
                      newdata = new_dat,
                      second.ord = F,
                      conf.level = 0.95, 
                      parm.type = "psi")

plot(x = seq(0, 5.6, length.out = 100), y = avg_prd$mod.avg.pred,
      xlab = 'Number People /100',
     ylab = 'Ocuupancy Probability', cex.axis = 1.5,
     cex.lab = 1.5, type= "l")
lines(x = seq(0, 5.6, length.out = 100),
      y = avg_prd$lower.CL,
      lwd = 2)
lines(x = seq(0, 5.6, length.out = 100),
      y = avg_prd$upper.CL,
      lwd = 2)

# calculating model averaged predictions detection 
avg_prd <- modavgPred(cand.set = cand.set,
                      newdata = new_dat,
                      second.ord = F,
                      conf.level = 0.95, 
                      parm.type = "detect")

plot(x = dist, y = avg_prd$mod.avg.pred,
     xlab = 'Proportion of area within 5km recently disturbed',
     ylab = 'Detection Probability', cex.axis = 1.5,
     cex.lab = 1.5, type= "l",
     ylim = c(0,0.1))
lines(x = dist,
      y = avg_prd$lower.CL,
      lwd = 2)
lines(x = dist,
      y = avg_prd$upper.CL,
      lwd = 2)


```

