---
title: "HW5"
author: "Peter Jenkins"
date: "4/1/2021"
output: html_document
---
1. Load data and place into an unmarkedFrameOccu object

```{r}
library(unmarked)
y <- read.csv("Bobcat.csv")
head(y)

bob_mat <- as.matrix(y)

site_cov <- read.csv('psi covariates.csv')

head(site_cov)

det_covs <- read.csv('p covariates.csv')
head(det_covs)

det_cov <- list(
  
  people = data.frame(det_covs))

nmix_data <- unmarkedFramePCount(y = bob_mat, siteCovs = site_cov,
                                 
                                 obsCovs = det_cov)


```

2. Fit an occupancy model that assumes conditional detection probability is a function of the number of people photographed at a site, and occupancy probability is a function of disturbance within 5km.
```{r}
fit <- pcount(formula = ~ people  ~ Dist_5km,
              data = nmix_data)

```


3. Interpret the effect of the number of people photographed at each site on conditional detection probability of bobcat. Verity your interpretation in R.
```{r}

summary(fit)

# log of expected count decreases for each 1 unit increase in proportion of area within 5km recently disturbed

# log estimated abundance 
e_y <- mean(as.matrix(y))
p <- plogis(coef(fit[2]))

e_y / p
exp(coef(fit[1]))


```

4. Re-create the Wald test for the effect of disturbance within 5km. What is the null hypothesis? Do you reject or fail to reject this null hypothesis?
```{r}
w <- -23.69 / 4.4307
w
## [1] -4.208314
2 * pnorm(-1 * abs(w))

```


We reject the null hypothesis because our p value is smaller than 0.05. 
  
5. Predict and plot the effect of disturbance within 5km on bobcat occupancy probability.
Do this over the range of observed disturbance within 5km.
```{r}

new_lam <- data.frame(Dist_5km = seq(min(site_cov$Dist_5km),max(site_cov$Dist_5km), length.out = 100))


prd <- predict(object = fit, newdata = new_lam,
               
               type = 'state')

prd


betas <- coef(fit)
betas

# linear model
exp(betas[1] + betas[2] * min(site_cov$Dist_5km) )


dist <- seq(min(site_cov$Dist_5km),max(site_cov$Dist_5km), length.out = 100)
# plotting 


mydata <- cbind(dist, prd)
# 2. Regression line + confidence intervals
library("ggplot2")
p <- ggplot(mydata, aes(dist, Predicted)) +
  geom_point() +
  xlab('Disturbance within 5 km') +# for the x axis label
ylab('Occupancy Probability')
  
# 3. Add prediction intervals
p + geom_line(aes(y = lower), color = "red", linetype = "dashed")+
  geom_line(aes(y = upper), color = "red", linetype = "dashed")

```
